Beef Pricing Dynamics in Canadian Supermarkets

### 2024-12-03 paper/paper.qmd
Add files via upload

The model is a Bayesian linear regression that predicts the current price of beef products. It uses the scaled old price old price as a numerical variable and includes vendor and month as categorical variables. This setup allows the model to account for the effects of historical prices, differences between supermarkets, and seasonal variations.

@@

- $\beta_0$ The intercept term, indicating the baseline price when all predictors are at their reference levels.
- $\beta_1$ The coefficient for old_price, representing the linear effect of the old price on the current price.
- $\beta_2$ The coefficient for vendor, representing the effect of different vendors on the current price (a categorical variable encoded as dummy variables).
- $\beta_3$ The coefficient for month, representing the effect of different months on the current price (a categorical variable encoded as dummy variables).

@@

A Bayesian approach was particularly suitable for this dataset, as it balances interpretability with predictive accuracy. By using priors, the model prevents overfitting and provides robust estimates even with limited data. This framework also enables us to quantify uncertainty in the predictions through posterior distributions, offering richer insights compared to traditional regression methods. Additionally, the inclusion of categorical variables (vendor and month) and a continuous predictor (scaled old prices) makes the model versatile for capturing both fixed and dynamic aspects of pricing. These attributes make Bayesian linear regression the optimal choice for this analysis.

>>>>

The model is a Bayesian linear regression that predicts the current price of beef products. It uses the month and old price as a numerical variable and includes vendor as categorical variables. This setup allows the model to account for the effects of historical prices, differences between supermarkets, and seasonal variations.

@@

- $\beta_0$ The intercept term, indicating the baseline price when all predictors are at their reference levels.
- $\beta_1$ The coefficient for old_price, representing the linear effect of the old price on the current price.
- $\beta_2$ The coefficient for vendor, representing the effect of different vendors on the current price (a categorical variable encoded as dummy variables).
- $\beta_3$ The coefficient for month, representing the effect of different months on the current price.

@@

A Bayesian approach was particularly suitable for this dataset, as it balances interpretability with predictive accuracy. By using priors, the model prevents overfitting and provides robust estimates even with limited data. This framework also enables us to quantify uncertainty in the predictions through posterior distributions, offering richer insights compared to traditional regression methods. Additionally, the inclusion of categorical variables (vendor) and a continuous predictor (old prices and month) makes the model versatile for capturing both fixed and dynamic aspects of pricing. These attributes make Bayesian linear regression the optimal choice for this analysis.
